Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-09 paper.tex
Some edits and comments

\usepackage{titlesec}
\titleformat{\chapter}
  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{hyperref} 
\usepackage{xcolor} 
\usepackage{ulem}

@@

@@

    \begin{abstract} Text of abstract.  
    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, 
        this data has been made widely available, however, accessibility to 
        such data remains limited due to infrastructural demands and lack of 
        performant software. Even with powerful infrastructure, such as High
        Performance Computing (HPC) infrastructure, software configured to 
        efficiently use resources is required to make execution time manageable
        . Ensuring efficient processing of data is 
        complex and limits accessibility to researchers with little programming 
        knowledge. Existing neuroimaging workflow engines primarily rely on HPC
        resource managers for performance and are therefore not well suited for
        BigData processing. As a result, frameworks for the efficient 
        processing of Big neuroimaging Data need to be developed\tristan{Add a sentence to transition from 'frameworks for efficient processing' to 'workflow engines'}

@@

        \tristan{Add a sentence to present the goal of the report.}
        This paper aims to evaluate the efficiency of current neuroimaging 
        frameworks in comparison to BigData frameworks and to determine the 
        extent to which BigData frameworks must be adapted to be effective
        neuroimaging workflow engines. 
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

        As we wish to evaluate and compare scientific workflow frameworks with
        Big Data frameworks for the processing of neuroimaging big data, we will
        break down our analyses by the four principal features necessary for
        effective neuroimaging workflow engines. The workflow engines discussed
        will include SPM, LONI, Nipype, PSOM, FastR, AA, Pydpiper 
        for neuroimaging; Taverna~\cite{doi:10.1093/bioinformatics/bth361}, 
        Pegasus, Make~\cite{10.3389/fninf.2016.00002}, 
        OpenMOLE~\cite{passerat2014openmole}, Nextflow~\cite{Di-Tommaso:2017aa}
        , Galaxy~\cite{Goecks2010}, 
        Toil for other scientific workflows; and Hadoop 
        MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, Spark and Dask for
        Big Data Frameworks. Chapter \ref{workcomp} will evaluate these 
        workflows on the basis of workflow composition: we will % 'in other words' is on my list of expressions to avoid
        examine the programming languages necessary to construct these workflows
        , as well as, how easy is it to add new tools. In Chapter 
        \ref{performance}, we will proceed to 
        investigate the strategies employed by these frameworks to improve the 
        performance of their workflows. Chapter \ref{portability} will discuss
        the portability of the workflows written using these frameworks. 
        Workflow reproducibility will be then discussed in Chapter 
        \ref{reproducibility}.

@@

            Neuroimaging big data comes in two formats: 1) large images and 2) 
            large datasets. Large images consist of singular images ranging 
            from 100s of gigabytes to terabytes in size. A well-known example 
            of 
            such an image is the BigBrain~\cite{Amunts1472}, a histological 
            image 
            of the brain of a healthy 69 year-old man. At its highest resolution
            , it is 1x1x20$\mu$m or 1TB in
            size. Other examples of large images can be found in electron 
            microscopy (EM)~\cite{Hildebrand:2017aa}, polarized light imaging 
            (PLI)~\cite{10.1007/978-3-319-12084-3_1} and micro computational
            tomography (microCT)~\cite{10.1371/journal.pone.0035691}. 
            Images at 
            such high resolution are important to researchers as they provide 
            insights into aspects not otherwise detectible in images at lower 
            resolutions. However, due to their size and lack of resources 
            available to process such images, research on this data remains 
            limited.

            In contrast, large datasets consist of many images acquired from
            multiple subjects, which are too large to be processed in their
            entirety. Notable examples of these datasets include the 
            Alzheimer's Disease Neuroimaging Initiative 
            (ADNI)~\cite{doi:10.1002/jmri.21049}, 
            OpenfMRI~\cite{POLDRACK2017259}, 
            the Autism Brain Imaging Data Exchange 
            (ABIDE)~\cite{Di-Martino:2013aa}, the Consortium for
            Reliability and Reproducibility (CoRR)~\cite{Zuo:2014aa}, the Human 
            Connectome Project (HCP)~\cite{VANESSEN201362} and the UK 
            Biobank~\cite{Miller:2016aa}, with the later
            two being flagship recent examples of large datasets in 
            neuroimaging.

@@

            OpenfMRI is a public repository originally conceived for the sharing
            of task-based fMRI data. Over time, it has grown to include EEG, MEG
            , resting-state fMRI (R-fMRI) and diffusion fMRI. Data made 
            available through
            OpenfMRI are obtained from both healthy and clinical individuals.
            Raw data must be made available for all datasets published to 
            OpenfMRI, however, preprocessed data may also accompany the raw data
            . At present, OpenfMRI houses data of a total 3372 subjects
            originating from 95 studies \note{obtained from website}.

@@

            All these initiative stress the importance of data-sharing and
            large-scale studies for the advancement of science. Many researchers
            are limited to the processing of subsets of the data available as 
            they lack compute and storage resources to process the data in its
            entirety. It is also important to note that many of the pipelines 
            developed by these initiatives made use of popular neuroimaging 
            toolboxes, such as FSL~\cite{SMITH2004S208}, FreeSurfer, ANTS, 
            CIVET, rather than
            creating their own. Therefore, there is a great need for Big Data
            frameworks that can leverage existing tools for the efficient 
            processing of large-scale datasets. Widespread availability and 
            accessibility of such frameworks will in turn lead to a greater
            volume of higher-impact studies.

@@

        \section{Infrastructure}\label{infrastructure}
            HPC infrastructure was initially designed for the processing of 
            compute-intensive applications. As a result, \tristan{I disagree with 'As a result' and I don't particularly agree with the previous sentence: even if
            resources aren't scarce, they are good reasons to optimize resource managers: energy consumption, application makespan, fairness, etc. Also,
            I'm not sure if resources are less scarce now than they used to be: with the rise of computational sciences, almost every researcher now
            needs a cluster, and Compute Canada (but also other infrastructures) has been very busy.}
            resource managers for these systems employ
            strategies to efficiently use the compute resources available for 
            processing. However, with the advent of big data, data-intensive 
            workflows are now beginning to become more present on HPC systems.
            Since the schedulers and workflow engines made for HPC systems are 
            'data-unaware', they are not efficient with the processing of Big
            Data

            HPC and Big Data infrastructure differ in many ways. We define HPC 
            systems as those that employ a shared parallel 
            filesystem (distinct from any compute node), such as Lustre, and 
            data-unaware schedulers. \tristan{and data-unaware schedulers?}.
            \tristan{You should say that this definition is 'by contrast to big data platforms' to explain where it's coming from.}
            The shared parallel filesystem is the slowest of
            all storage devices found on HPC as it is shared amongst all users 
            of the system.
            Compute nodes in such infrastructure may not contain any local 
            storage. However, network interconnects between nodes are typically 
            fast.
            Resources managers used in HPC infrastructures typically employ 
            batch scheduling and execute heterogenous workflows. 
            Examples of schedulers used in HPC clusters 
            include SLURM~\cite{yoo2003slurm}, 
            PBS~\cite{10.1007/3-540-60153-8_34}, SGE~\cite{SGE} and 
            HTCondor~\cite{htcondor}.

            Big Data infrastructure \tristan{Call it 'Big Data infrastructure?'} differs from HPC infrastructures in that
            it uses less-costly hardware. It follows a shared-nothing paradigm
            where the storage is not shared between nodes. That is, all compute
            nodes are responsible for storing the data as well. File systems
            used within such infrastructure are distributed filesystems, such 
            as the
            Hadoop Distributed File System (HDFS)~\cite{5496972} and 
            Alluxio~\cite{Li:2014:TRM:2670979.2670985}.
            Fault-tolerance of the data is achieved through replication across
            nodes. As transferring data across the network is typically costly 
            in such infrastructures, frameworks designed to run on Big Data 
            infrastructure attempt to limit the transfer of data across nodes. 
            Resource managers created specifically for these systems
            typically expect dedicated infrastructure, as is the case with 
            Yet Another Resource Negotiator 
            (YARN)~\cite{Vavilapalli:2013:AHY:2523616.2523633}, but can also 
            execute heterogeneous workflows, as can be seen with 
            Mesos~\cite{hindman2011mesos}.

@@

        Workflows, also known as pipelines, describe the flow of data from one 
        activity to another and are typically represented by a graph. \tristan{Not necessarily a DAG, see Nipype for instance, it has loops.
        I don't think you want to enter in a taxonomy of languages though, just say 'graph'}. Workflow activities, also referred to as 
        workflow interfaces or stages, encapsulate a tool through the use of an
        interface\tristan{Mention 'interface' here as an intro to sub-sec 2}. 
        Tools can be represented in many forms, including: command-line tools, 
        custom scripts integrated directly into the workflow code, web 
        services and software libraries.\tristan{A script is a command-line tool. I think tools may also be web services (Taverna) or just 
        software libraries (Dask)}. A workflow activity with all its 
        dependencies is considered ready to be staged.

        For a workflow engine to be considered successful, it must be simple to 
        compose workflows using that framework. New researchers should be able 
        to acquire the skills to compose worfklows without
        investing a significant amount of time in understanding it. Reuse of
        existing tools is also common occurrence in neuroinformatics, and 
        workflow engines need to accomodate this by undertaking a modular 
        approach and being easily extensible. With the sharing of data also 
        comes the sharing of workflows. For labs to share their workflows with
        the community at large, these workflows must be easily understood by the
        community. This relates to the reproducibility of studies, as it is 
        more than likely that easily understood workflows are reproducible. 
        \tristan{Perhaps mention that this also relates to the sharing
        of reproducible results -- VH: not sure if this is what you meant}.

@@

        \section{Workflow language}\label{lang}
            The choice of workflow language (e.g. GUI, scripting language,
            programming language) greatly
            affects how easily it will get picked up by the community. Complex
            languages may mean greater efficiency of engines
            , but also result in extensive time spent on learning and developing
            the pipelines. Moreover, some researchers may not be well-versed in
            programming and may require very easy-to-use engines.

            To address this issue, some workflow engines have taken the role
            of employing a graphical user interface (GUI). This makes workflow 
            composition accessible to all researchers, including those with
            little to no programming background. The caveat to employing a 
            workflow engine with a GUI interface for workflow composition is
            that users are greatly limited by the tasks they can perform. For 
            instance, workflows that perform any form of iterative operations
            (e.g. machine-learning workflows) or workflows with conditionals
            (e.g. given a certain input, do not execute this activity) cannot
            be represented by a GUI. 
            Workflow engines that employ a worfklow composition GUI include:
            SPM, Taverna, OpenMOLE, LONI and Galaxy.
            \tristan{My conclusion on GUIs is that they are suitable only
            for very simple pipelines. Perhaps the intro could give an idea
            of the types of pipelines encountered, showing a simple one and
            a more complex one. No way fmriprep is described in a GUI.}

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI and OpenMOLE, allow more 
            experienced users to script their workflow. Four different 
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and GNU/Octave \tristan{I think the official name is GNU/Octave}, 3) Shell Scripting languages \tristan{scripting languages}, 4) Domain Specific Language (DSL). \tristan{I don't see why CWL and XML don't qualify as DSL} Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus, Toil,
            MapReduce, Dask and Spark. \tristan{Dask? Spark?}.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python include: a wide array of available libraries,
            particularly neuroimaging-specific ones such as Nilearn~\cite{nilearn} and Nibabel~\cite{matthew_brett_2018_1287921}
            and more general, but heavily used in neuroimaging, libraries such
            as SciPy~\cite{scipy} and Scikit Learn~\cite{pedregosa2011scikit} \tristan{mention scipy, sci-kit learn, etc}
            , and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python
            , whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

            Another widely-used high-level programming language in the 
            neuroinformatics community is MATLAB/Octave. Workflow engines which
            use this language include: PSOM, SPM, and Automatic analysis. Like 
            Python, MATLAB and GNU/Octave are platform-independent and also easy to use. In addition, MATLAB has 
            been found to exhibit better performance compared to Python for 
            common technical tasks. It also provides an integrated environment
            for programming. A 
            downfall to imposing MATLAB as programming language is that it is proprietary, thus imposing a fee on 
            anyone wishes to program in that language or have access to their diverse array of libraries \tristan{You should also mention the
            advantage of Matlab: easy to learn, integrated ecosystem, 
            performance, see 
            \url{https://www.mathworks.com/products/matlab/matlab-vs-python.html}}. 
            To offset this, PSOM ensure compatibility with the GNU 
            Octave programming language, thus 
            removing the requirement to purchase software for workflow 
            composition.

            \tristan{This paragraph should start with ``Shell scripting ...". The structuration of this section by language
            is great, but the writing of this paragram doesn't put it forward.}
            Many researchers have been exposed to Shell scripting languages as 
            several scientific workflows are composed in UNIX-based 
            environments. These environments typically have Make installed by
            default eliminating the need for users to install it manually.
            These features make workflow engine Make desirable for scientific
            computing.
            As Make is exclusive to UNIX environments, it is not 
            platform-independent. However, Make has the flexibility of being
            easily extensible and can interface directly with any command-line
            tool. Shell scripting, nevertheless, is generally not as intuitive
            as programming languages like Python and MATLAB.

            Domain Specific Languages (DSL) are also used to compose workflows.
            The can be made up of any language (e.g. XML, CWL, Scala and Groovy) that has been adapted for use in 
            a specific domain.
            XML is the language of choice in  
            workflow engines such as Pegasus and LONI. Using a markup language
            such as XML have the advantage of having access to 
            language-independent parsers and validators. However, the use of a 
            markup language removes some of the flexibility. \tristan{Before you explain the downside, 
            you should say what are the advantages: mainly language-independent parsers and validation.} Users wanting to 
            include 
            custom scripts as workflow activities must do so by creating a 
            command-line interface for their custom script. Pegasus, however,
            does provide many interfaces to describe the pipeline in including
            Java, Python and Perl.

            Another example of a DSL is a Common Workflow
            Language (CWL). CWL can be seen in workflow engines like Toil. 
            It is similar to XML in that its main difference is that it is 
            YAML/JSON-based.
            \tristan{It is a YAML-based DSL for workflows. Just like Pegasus' DAX is an XML-based DSL}. Files can either be written in JSON or YAML formats. CWL is 
            based on workflow engines like Make \tristan{? I don't think CWL relies on make, does it?}, but differs in that tasks
            are isolated \tristan{inputs and outputs are somewhat explicit in a Makefile}. Similarly to 
            make, CWL is only available on Linux and OSX platforms \tristan{strange, the CWL interpreter is in Python so I'd expect it to run anywhere}. \note{cite cwl 
            website}

            Programming language-based DSL workflow engines include OpenMOLE, which is  
            derived from Scala, and Nextflow \tristan{Nextflow?}, which is derived from the Apache 
            Groovy programming language. While both are written on top of the 
            Java platform, Nextflow is only compatible with Linux and OSX \tristan{How about using Linux everywhere instead of UNIX-based?} 
            environments. These programming languages are not otherwise commonly
            used in neuroimaging, meaning that users must learn specific 
            languages that would not otherwise be used in their field.

            With the exeption of LONI, neuroimaging workflows tend to stick to
            commonly used programming languages such as Python and MATLAB/Octave
            . This is likely due to the fact that the learning curve to adapt to
            these languages is relatively low and users can harness the 
            flexibility of these languages.

@@

                The data structures made available by workflow engines dictate
                the maximum performance available to a workflow. Scientific
                workflows are generally file-based, meaning that data, in the 
                form of files, will be 
                transferred across the network and I/O must be
                performed at each workflow activity. It is possible for some 
                data to remain in memory with scientific workflows, but these 
                are typically basic types such as numbers, strings and lists.
                More complex types, however, can be seen in workflow engines 
                such as PSOM which accept nested structures as long as the
                structures's terminal fields are either strings or cells of 
                strings. Due to workflow engines being file-based, performance
                is limited, particularly with BigData.

@@

                The MapReduce key-value pair data structure is an effective 
                data structure for describing workflows, but is limited to only
                batch workflows and can be improved in terms of performance. 
                The resilient distributed datasets 
                (RDD)~\cite{Zaharia:2012:RDD:2228298.2228301}, Spark's main
                abstractions addresses these 
                issues by being a shared-memory structure. They 
                are immutable and only permit coarse-grained edits to ensure 
                fault-tolerance. The RDDs shared memory enables a reduction in
                I/O operation as data can be maintained in memory. Spark will only read 
                and write data to and from disk when data exceeds available 
                memory or when data needs to be shuffled around, as is the case
                with all Spark actions and certain transformations such as 
                \textit{join} and \textit{reduceByKey}. Moreover, RDDs permit
                workflows to be evaluated lazily. By being able to look at the
                entire workflow before evaluation, Spark can select the most
                performant methods for evaluating the data. 
                The RDD additionally expands its use to additional applications
                by enabling the expression of both iterative and interactive
                workflows. Furthermore, Spark provides an additional data 
                structure to enable the efficient processing of tabular data.
                This data structure is know as the 
                DataFrame~\cite{Armbrust:2015:SSR:2723372.2742797}.

>>>>

\usepackage{titlesec}
\titleformat{\chapter}
  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{hyperref} 
\usepackage{xcolor} 
\usepackage{ulem}                     
\usepackage{pdfpages} 
\usepackage{rotating}

@@

% Fix link colors
\hypersetup{
    colorlinks = true,
    linkcolor=red,
    citecolor=red,
    urlcolor=blue,
    linktocpage % so that page numbers are clickable in toc
}
\setcounter{tocdepth}{1}

@@

    \begin{abstract} Text of abstract.  
    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, 
        this data has been made widely available, however, accessibility to 
        such data remains limited due to infrastructural demands and lack of 
        performant software. Even with powerful infrastructure, such as High
        Performance Computing (HPC) infrastructure, software configured to 
        efficiently use resources is required to make execution time manageable
        . Ensuring efficient processing of data is 
        complex and limits accessibility to researchers with little programming 
        knowledge. Existing neuroimaging workflow engines primarily rely on HPC
        resource managers for performance and are therefore not well suited for
        BigData processing. As a result, frameworks for the efficient 
        processing of Big neuroimaging Data need to be developed.

@@

        This paper aims to evaluate the efficiency of current neuroimaging 
        frameworks in comparison to BigData frameworks and to determine the 
        extent to which BigData frameworks must be adapted to be effective
        neuroimaging workflow engines. 
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

        As we wish to evaluate and compare scientific workflow frameworks with
        Big Data frameworks for the processing of neuroimaging big data, we will
        break down our analyses by the four principal features necessary for
        effective neuroimaging workflow engines. The workflow engines discussed
        will include SPM, LONI, Nipype, PSOM, FastR, AA, Pydpiper 
        for neuroimaging; Taverna~\cite{doi:10.1093/bioinformatics/bth361}, 
        Pegasus, Make~\cite{10.3389/fninf.2016.00002}, 
        OpenMOLE~\cite{passerat2014openmole}, Nextflow~\cite{Di-Tommaso:2017aa}
        , Galaxy~\cite{Goecks2010}, 
        Toil for other scientific workflows; and Hadoop 
        MapReduce~\cite{Dean:2008:MSD:1327452.1327492}, Spark and Dask for
        Big Data Frameworks. Chapter \ref{workcomp} will evaluate these 
        workflows on the basis of workflow composition: we will % 'in other words' is on my list of expressions to avoid
        examine the programming languages necessary to construct these workflows,
        as well as, how easy is it to add new tools. In Chapter 
        \ref{performance}, we will proceed to 
        investigate the strategies employed by these frameworks to improve the 
        performance of their workflows. Chapter \ref{portability} will discuss
        the portability of the workflows written using these frameworks. 
        Workflow reproducibility will be then discussed in Chapter 
        \ref{reproducibility}.

@@

            Neuroimaging big data comes in two formats: 1) large images and 2) 
            large datasets. Large images consist of singular images ranging 
            from 100s of gigabytes to terabytes in size. A well-known example 
            of 
            such an image is the BigBrain~\cite{Amunts1472}, a histological 
            image 
            of the brain of a healthy 69 year-old man. At its highest resolution,
            it is 1x1x20$\mu$m or 1TB in
            size. Other examples of large images can be found in electron 
            microscopy (EM)~\cite{Hildebrand:2017aa}, polarized light imaging 
            (PLI)~\cite{10.1007/978-3-319-12084-3_1} and micro computational
            tomography (microCT)~\cite{10.1371/journal.pone.0035691}. 
            Images at 
            such high resolution are important to researchers as they provide 
            insights into aspects not otherwise detectible in images at lower 
            resolutions. However, due to their size and lack of resources 
            available to process such images, research on this data remains 
            limited.

            \tristan{It could be useful to refer to this special issue on workflows
            in neuroimaging in intro:
            https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2844796/}
            \tristan{This special issue on data sharing is also worth citing:
            https://www.sciencedirect.com/journal/neuroimage/vol/124/part/PB}
            In contrast, large datasets consist of many images acquired from
            multiple subjects, which are too large to be processed in their
            entirety. Notable examples of these datasets include the 
            Alzheimer's Disease Neuroimaging Initiative 
            (ADNI)~\cite{doi:10.1002/jmri.21049}, 
            OpenfMRI~\cite{POLDRACK2017259}, 
            the Autism Brain Imaging Data Exchange 
            (ABIDE)~\cite{Di-Martino:2013aa}, the Consortium for
            Reliability and Reproducibility (CoRR)~\cite{Zuo:2014aa}, the Human 
            Connectome Project (HCP)~\cite{VANESSEN201362} and the UK 
            Biobank~\cite{Miller:2016aa}, with the later
            two being flagship recent examples of large datasets in 
            neuroimaging.

@@

            OpenfMRI is a public repository originally conceived for the sharing
            of task-based fMRI data. Over time, it has grown to include EEG, MEG,
            resting-state fMRI (R-fMRI) and diffusion fMRI. Data made 
            available through
            OpenfMRI are obtained from both healthy and clinical individuals.
            Raw data must be made available for all datasets published to 
            OpenfMRI, however, preprocessed data may also accompany the raw data
            . At present, OpenfMRI houses data of a total 3372 subjects
            originating from 95 studies \note{obtained from website}.

@@

            All these initiative stress the importance of data-sharing and
            large-scale studies for the advancement of science. Many researchers
            are limited to the processing of subsets of the data available as 
            they lack compute and storage resources to process the data in its
            entirety. It is also important to note that many of the pipelines 
            developed by these initiatives made use of popular neuroimaging 
            toolboxes, such as FSL~\cite{SMITH2004S208} \tristan{You should cite it the first time you mention it}, FreeSurfer, ANTS, 
            CIVET, rather than
            creating their own. Therefore, there is a great need for Big Data
            frameworks that can leverage existing tools for the efficient 
            processing of large-scale datasets. Widespread availability and 
            accessibility of such frameworks will in turn lead to a greater
            volume of higher-impact studies.

@@

        \section{Infrastructure}\label{infrastructure}
            HPC infrastructure was initially designed for the processing of 
            compute-intensive applications. As a result,
            resource managers for these systems employ
            strategies to efficiently use the compute resources available for 
            processing. However, with the advent of big data, data-intensive 
            workflows are now beginning to become more present on HPC systems.
            Since the schedulers and workflow engines made for HPC systems are 
            'data-unaware', they are not efficient with the processing of Big
            Data

            HPC and Big Data infrastructure differ in many ways. We define HPC 
            systems as those that employ a shared parallel 
            filesystem (distinct from any compute node), such as Lustre, and 
            data-unaware schedulers.
            The shared parallel filesystem is the slowest of
            all storage devices found on HPC as it is shared amongst all users 
            of the system.
            Compute nodes in such infrastructure may not contain any local 
            storage. However, network interconnects between nodes are typically 
            fast.
            Resources managers used in HPC infrastructures typically employ 
            batch scheduling and execute heterogenous workflows. 
            Examples of schedulers used in HPC clusters 
            include SLURM~\cite{yoo2003slurm}, 
            PBS~\cite{10.1007/3-540-60153-8_34}, SGE~\cite{SGE} and 
            HTCondor~\cite{htcondor}.

            Big Data infrastructure differs from HPC infrastructures in that
            it uses less-costly hardware. It follows a shared-nothing paradigm
            where the storage is not shared between nodes. That is, all compute
            nodes are responsible for storing the data as well. File systems
            used within such infrastructure are distributed filesystems, such 
            as the
            Hadoop Distributed File System (HDFS)~\cite{5496972} and 
            Alluxio~\cite{Li:2014:TRM:2670979.2670985}.
            Fault-tolerance of the data is achieved through replication across
            nodes. As transferring data across the network is typically costly 
            in such infrastructures, frameworks designed to run on Big Data 
            infrastructure attempt to limit the transfer of data across nodes. 
            Resource managers created specifically for these systems
            typically expect dedicated infrastructure, as is the case with 
            Yet Another Resource Negotiator 
            (YARN)~\cite{Vavilapalli:2013:AHY:2523616.2523633}, but can also 
            execute heterogeneous workflows, as can be seen with 
            Mesos~\cite{hindman2011mesos}.

@@

        Workflows, also known as pipelines, describe the flow of data from one 
        activity to another and are typically represented by a graph. Workflow activities, also referred to as 
        workflow interfaces or stages, encapsulate a tool through the 
        use of an interface. Tools can be represented in many forms, 
        including: command-line tools, custom scripts integrated 
        directly into the workflow code, web services and software 
        libraries. A workflow activity with all its 
        dependencies is considered ready to be staged.

        For a workflow engine to be considered successful, it must be simple to 
        compose workflows using that framework. New researchers should be able 
        to acquire the skills to compose worfklows without
        investing a significant amount of time in understanding it. Reuse of
        existing tools is also common occurrence in neuroinformatics, and 
        workflow engines need to accomodate this by undertaking a modular 
        approach and being easily extensible. With the sharing of data also 
        comes the sharing of workflows. For labs to share their workflows with
        the community at large, these workflows must be easily understood by the
        community. This relates to the reproducibility of studies, as it is 
        more than likely that easily understood workflows are reproducible. 
        \tristan{Perhaps mention that this also relates to the sharing
        of reproducible results -- VH: not sure if this is what you meant --
        yeah, we might want to explicit this further in the paper but
        good enough for now}.

@@

        \section{Workflow language}\label{lang}
            The choice of workflow language (e.g. GUI, scripting language,
            programming language) greatly
            affects how easily it will get picked up by the community. Complex
            languages may mean greater efficiency of engines,
            but also result in extensive time spent on learning and developing
            the pipelines. Moreover, some researchers may not be well-versed in
            programming and may require very easy-to-use engines.

            To address this issue, some workflow engines have taken the role
            of employing a graphical user interface (GUI). This makes workflow 
            composition accessible to all researchers, including those with
            little to no programming background. The caveat to employing a 
            workflow engine with a GUI interface for workflow composition is
            that users are greatly limited by the tasks they can perform. For 
            instance, workflows that perform any form of iterative operations
            (e.g. machine-learning workflows) or workflows with conditionals
            (e.g. given a certain input, do not execute this activity) can hardly
            be represented by a GUI. 
            Workflow engines that employ a worfklow composition GUI include:
            SPM, Taverna, OpenMOLE, LONI and Galaxy.

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI and OpenMOLE, allow more 
            experienced users to script their workflow. Four different 
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and GNU/Octave, 3) Shell Scripting languages, 4) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus, Toil,
            MapReduce, Dask and Spark.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python include: a wide array of available libraries,
            particularly neuroimaging-specific ones such as Nilearn~\cite{nilearn} and Nibabel~\cite{matthew_brett_2018_1287921}
            and more general, but heavily used in neuroimaging, libraries such
            as SciPy~\cite{scipy} and Scikit Learn~\cite{pedregosa2011scikit},
             and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python,
            whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

            Another widely-used high-level programming language in the 
            neuroinformatics community is MATLAB/Octave. Workflow engines which
            use this language include: PSOM, SPM, and Automatic analysis. Like 
            Python, MATLAB and GNU/Octave are platform-independent and also easy to use. In addition, MATLAB has 
            been found to exhibit better performance compared to Python for 
            common technical tasks. It also provides an integrated environment
            for programming. A 
            downfall to imposing MATLAB as programming language is that it is proprietary, thus imposing a fee on 
            anyone wishes to program in that language or have access to their diverse array of libraries. 
            To offset this, PSOM ensure compatibility with the GNU 
            Octave programming language, thus 
            removing the requirement to purchase software for workflow 
            composition.

            Many researchers have been exposed to Shell scripting languages as 
            several scientific workflows are composed in UNIX-based 
            environments. These environments typically have Make installed by
            default eliminating the need for users to install it manually.
            These features make workflow engine Make desirable for scientific
            computing.
            As Make is exclusive to UNIX environments, it is not 
            platform-independent. However, Make has the flexibility of being
            easily extensible and can interface directly with any command-line
            tool. Shell scripting, nevertheless, is generally not as intuitive
            as programming languages like Python and MATLAB.

            Domain Specific Languages (DSL) are also used to compose workflows.
            They can be made up of any language (e.g. XML, Scala and Groovy) that has been adapted for use in 
            a specific domain.
            XML is the language of choice in  
            workflow engines such as Pegasus and LONI. Using a markup language
            such as XML has the advantage of having access to 
            language-independent parsers and validators. However, the use of a 
            markup language removes some of the flexibility. Users wanting to 
            include 
            custom scripts as workflow activities must do so by creating a 
            command-line interface for their custom script. Pegasus, however,
            does provide many interfaces to describe the pipeline in including
            Java, Python and Perl.

            Another example of a DSL is the Common Workflow
            Language (CWL). CWL can be seen in workflow engines like Toil. 
            It is similar to XML in that its main difference is that it is 
            YAML/JSON-based. \tristan{Review the previous sentence, it doesn't make sense. And the end of the paragraph is till fuzzy too}
            \tristan{It is a YAML-based DSL for workflows. Just like Pegasus' DAX is an XML-based DSL}. Files can either be written in JSON or YAML formats. CWL is 
            based on workflow engines like Make \tristan{? I don't think CWL relies on make, does it?}, but differs in that tasks
            are isolated \tristan{inputs and outputs are somewhat explicit in a Makefile}. Similarly to 
            make, CWL is only available on Linux and OSX platforms \tristan{strange, the CWL interpreter is in Python so I'd expect it to run anywhere}. \note{cite cwl 
            website}

            Programming language-based DSL workflow engines include OpenMOLE, which is  
            derived from Scala, and Nextflow, which is derived from the Apache 
            Groovy programming language. While both are written on top of the 
            Java platform, Nextflow is only compatible with Linux and OSX \tristan{How about using Linux everywhere instead of UNIX-based? You still have UNIX here and there.} 
            environments. These programming languages are not otherwise commonly
            used in neuroimaging, meaning that users must learn specific 
            languages that would not otherwise be used in their field.

            With the exeption of LONI, neuroimaging workflows tend to stick to
            commonly used programming languages such as Python and MATLAB/Octave.
            This is likely due to the fact that the learning curve to adapt to
            these languages is relatively low and users can harness the 
            flexibility of these languages.

@@

                The data structures made available by workflow engines dictate
                the maximum performance available to a workflow. \tristan{I don't understand the previous sentence, 
                in particular the 'maximum' idea. You should review that.} Scientific
                workflows are generally file-based, meaning that data, in the 
                form of files, will be 
                transferred across the network and I/O must be
                performed at each workflow activity. It is possible for some 
                data to remain in memory with scientific workflows, but these 
                are typically basic types such as numbers, strings and lists.
                More complex types, however, can be seen in workflow engines 
                such as PSOM which accept nested structures as long as the
                structures's terminal fields are either strings or cells of 
                strings. Due to workflow engines being file-based, performance
                is limited, particularly with BigData.

@@

                The MapReduce key-value pair data structure is an effective 
                data structure for describing workflows, but is limited to only
                batch workflows and can be improved in terms of performance. 
                The resilient distributed datasets 
                (RDD)~\cite{Zaharia:2012:RDD:2228298.2228301}, Spark's main
                abstractions addresses these 
                issues by being a shared-memory structure. They 
                are immutable and only permit coarse-grained edits, to ensure 
                fault-tolerance. The RDDs shared memory enables a reduction in
                I/O operation as data can be maintained in memory. Spark will only read 
                and write data to and from disk when data exceeds available 
                memory or when data needs to be shuffled around, as is the case
                with all Spark actions and certain transformations such as 
                \textit{join} and \textit{reduceByKey}. Moreover, RDDs permit
                workflows to be evaluated lazily. By being able to look at the
                entire workflow before evaluation, Spark can select the most
                performant methods for evaluating the data. 
                The RDD additionally expands its use to additional applications
                by enabling the expression of both iterative and interactive
                workflows. Furthermore, Spark provides an additional data 
                structure to enable the efficient processing of tabular data.
                This data structure is know as the 
                DataFrame~\cite{Armbrust:2015:SSR:2723372.2742797}.
